---
title: 'STA6235: Project 1 Assignment'
author: "Gabriel Rivera Alvarez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Libraries Used
```{r}
# Libraries to be used
library(tidyverse)
```

#### Dataset Reading
```{r}
# Load dataset
data <- read.csv("/Users/gabrielrivera/Documents/Maestria/Courses/4- Summer 2022/STA6235/Project 1/Walmart_Store_sales.csv")
```

#### Dataset Display
```{r}
#Display first 5 rows of the dataset
head(data,5)
```

#### Dataset Extraction and Display
```{r}
# Put variables of interest as a tibble
dataset <- tibble(Temp = data$Temperature, WkSales = data$Weekly_Sales)
head(dataset,5)
```

#### Dataset Description
The dataset being used comes from Kaggle. The link to the dataset is: https://www.kaggle.com/datasets/rutuspatel/walmart-dataset-retail. This dataset contains columns that are: Store, Date, Weekly_Sales, Holiday_Flag, Temperature, Fuel_Price, CPI, and Unemployment. From this dataset, I will be using Temperature and Weekly_Sales. I will use Temperature as my predictor and Weekly_Sales as my outcome. I want to see if the Temperature can be a good predictor for the weekly sales. The whole data is spread across 6435 observations over 8 columns or variables.

#### Statistical Summary
```{r}
# Statistics summary
means <- summarize(dataset, mean_Temp = mean(dataset$Temp, na.rm=TRUE), sd_Temp =sd(dataset$Temp, na.rm=TRUE),
                            mean_WkSales = mean(dataset$WkSales, na.rm=TRUE), sd_WkSales =sd(dataset$WkSales, na.rm=TRUE))
means
```

Doing a statistics summary we get an average temp of 60.66378 degrees with a standard deviation of 18.44493. for weekly sales we have an average of $1,046,965	with a standard deviation of 564,366.6.

#### Model Construction
```{r}
# Model Construction
model <- lm(WkSales ~ Temp, data = dataset)
model_summary <- summary(model)
model_summary
```

#### Intercept and Slope Extraction
```{r}
# Coefficient Extraction
lin_coeff <- coefficients(model)
lin_coeff
```

#### Test Statistics Extraction
```{r}
# Get Test Statistics
t0 <- as_tibble(model_summary[[4]])
t0
```

#### Resulting linear model is:

\[ \hat{WkSales} = `r round(lin_coeff[1],2)` + `r round(lin_coeff[2],2)`*Temp \]

#### Construction of a 95% Confidence Interval
```{r}
# 95% Confidence Interval Construction
model_ci <- as_tibble(confint(model, level=0.95))
model_ci
```
Confidence Interval for the Intercept:

- $`r round(model_ci$"2.5 %"[1], digits = 2)` \le \beta_0 \le `r round(model_ci$"97.5 %"[1], digits = 2)`$

Confidence Interval for the Slope:

- $`r round(model_ci$"2.5 %"[2], digits = 2)` \le \beta_1 \le `r round(model_ci$"97.5 %"[2], digits = 2)`$

#### Hypothesis Testing

Hypotheses:

- $H_0: \ \beta_1 = 0$
- $H_1: \ \beta_0 \ne 0$

Test Statistic:

- $t_0 = `r round(t0$"t value"[2], digits = 2)`$

$p$-Value:

- $p =`r t0$"Pr(>|t|)"[2]`$

Rejection Region:

- Reject $H_0$ if $p < \alpha$ for $\alpha = 0.05$

Conclusions/Interpretation:

- Reject $H_0$. There is sufficient evidence to suggest the temperature predicts the weekly sales.

#### $R^2$ of the resulting model:

The $R^2$ is:

- $R^2 = `r round(model_summary$r.squared, digits=5)`$

The $R^2 = `r round(model_summary$r.squared, digits=5)`$ means the regression line is accounting for $`r round(model_summary$r.squared, digits=5)*100`$% of the variability in the weekly sales outcome variable.

#### Data Visualization

Scatter plot of WkSales as function of Temp:

```{r}
dataset %>% ggplot(aes(x=Temp, y=WkSales))+
  geom_point(size=0.1) +
  geom_smooth(method=lm, se=FALSE, formula=y~x) +
  theme_minimal()
```

#### Final Remarks:

Based on model results it can be said that temperature predicts weekly sales. That being said, this model only takes into consideration one predictor and the dataset contains multiple columns that can help to develop a better model. Also, the dataset is taken across 45 different stores that can be located in different regions, thus temperature could affect differently in each store. Taking a look at a scatter plot of the data, the linear regression found does not seem like a good fit. The line is close to being a horizontal line. The model developed can only explain approximately 0.4% of the variability of the weekly sales outcome variable, which seems too little for a good linear fit. In conclusion, although it was found through the hypothesis testing that temperature predicts weekly sales, taking into consideration other factors like visualization and $R^2$ it can be seen that the model is not a good linear fit.